---
title: "Income Along the Boston T [Orange Line]"
author: "Jorge A Rodriguez MD"
date: '2017-12-12'
slug: income-along-the-boston-t
---

As a native New Yorker, I was recently intrigued by a visualization of household income in NY to highlight inequality on the [New Yorker](https://projects.newyorker.com/story/subway/). Having live in Boston for going on 9 years, I've started to call *this place* home. Could I make a similar representation of income in Boston? **That was my challenge** 

### PART I: <span style="color:orange">The Orange Line</span>

To start I used these packages:  
1. [acs](https://cran.r-project.org/web/packages/acs/index.html): A simple way to pull census data  
2. [httr](https://cran.r-project.org/web/packages/httr/index.html): Let's your work with URLs  
3. [tidyverse](https://www.tidyverse.org/): *THE* R package  
4. [plotly](https://plot.ly/r/): Interactive graphing in R
```{r echo=TRUE, eval=TRUE, message=FALSE}
library(acs)
library(httr)
library(plotly)
library(tidyverse)
```

I pulled together the names and locations(longitide/latitude) of the [Orange Line stations](../mbta_orange.csv). I needed to change the longitude and latitude to numerics.
```{r eval=FALSE}
mbta = read.csv("mbta_orange.csv")
mbta$Longitude = as.numeric(mbta$Longitude)
mbta$Latitude = as.numeric(mbta$Latitude)
```

Next up, creating some empty columns that we will fill in down below. 
```{r eval=FALSE}
mbta$tract = NaN
mbta$county = NaN
mbta$income = NaN
```

#### Census Tracts
Since I wanted to know the median household income by census tract, I needed to convert the coordinates I had to census tract. I used the httr package to use the census geocoder website. The end of the code gets a little confusing due to the way the geocoding website returns the data.
```{r echo=TRUE, eval=FALSE}
for(i in 1:nrow(mbta)) {
  x= as.character(mbta$Longitude[i])
  y = as.character(mbta$Latitude[i])
  url = sprintf("https://geocoding.geo.census.gov/geocoder/geographies/coordinates?x=%s&y=%s&benchmark=4&vintage=4", x, y)
  print(url)
  r= GET(url)
  g = content(r, "parsed")
  print(i)
  mbta$tract[i]= as.character(g$result$geographies$`Census Tracts`[[1]][17]$TRACT[[1]])
  mbta$county[i] = as.character(g$result$geographies$Counties[[1]]$NAME[[1]])
}
```

#### Median Household Income
With the census tracts for each Orange line T stop in hand, I used the acs package to get income data. I created a function to do just that, then used the mapply function to repeat it for each station.
```{r eval=FALSE}
tractIncome = function(tract,county) {
  my.tract=geo.make(state="MA", county=county, tract=tract)
  median.income=acs.fetch(geo=my.tract, endyear=2015,  table.name ="B19013")
  as.numeric(estimate(median.income[1,1]))[[1]]
}

mbta$income = mapply(tractIncome, mbta$tract, mbta$county)
```

```{r eval=TRUE, echo=FALSE}
mbta=read.csv("mbta_full_orange.csv")
```

#### Visualization
After a few trials, I realized I needed to change the factor levels of the station names to display them in the order of the stops. With that sorted out, it was on to using plotly to create the interactive graph. *Note that the plotly function has an underscore.* This took me a while to figure out. So there you have it! Median household income along Boston's Orange line
```{r eval=TRUE, error=FALSE, message=FALSE}
mbta$Station_Name = as.character(mbta$Station_Name)
mbta$Station_Name = factor(mbta$Station_Name, levels=unique(mbta$Station_Name))

plot_ly(mbta, x=~Station_Name, hoverinfo="none") %>%
  add_trace(y=~Income, type="scatter", mode="line", line = list(color="orange", width=6)) %>%
  add_trace(y=~Income, type="scatter", mode="markers", marker=list(color="black", size=8), hoverinfo= 'text',
        text = ~paste('</br><b>Station Name:</b>', Station_Name,
                      '</br> Income:', Income,
                      '</br> Tract:', tract)) %>%
  layout(title="Median Income Along the T", showlegend=FALSE, xaxis=list(title="", tickangle=-90), yaxis=list(title="Median Household Income", range=c(0, 200000)), margin=list(b=160))

```
*Note: Census tracts cover about 1,200 and 8,000 people*