---
title: Income Along the Boston T [Red Line]
author: Jorge A Rodriguez MD
date: '2017-12-21'
slug: income-along-the-boston-t-ii
---
In the [first part](http://translatedmedicine.com/post/income-along-the-boston-t/) of this series, I went through my approach for creating a [New Yorker](https://projects.newyorker.com/story/subway/) inspired visualization of income along <span style="color:orange">The Orange Line</span> on the Boston T. 

For this post, we take a look at the <span style="color:red">The Red Line</span>. Below you will see median household income along both lines of the Boston Red Line. 

####Ashmont/Mattapan Red Line
```{r eval = TRUE, echo= FALSE, message=FALSE}
library(plotly)
mbta = read.csv("mbta_full_red_mattapan.csv")
```

```{r eval=TRUE, error=FALSE, message=FALSE, echo=FALSE}
mbta$Station_Name = as.character(mbta$Station_Name)
mbta$Station_Name = factor(mbta$Station_Name, levels=unique(mbta$Station_Name))

plot_ly(mbta, x=~Station_Name, hoverinfo="none") %>%
  add_trace(y=~Income, type="scatter", mode="line", line = list(color="red", width=6)) %>%
  add_trace(y=~Income, type="scatter", mode="markers", marker=list(color="black", size=8), hoverinfo= 'text',
        text = ~paste('</br><b>Station Name:</b>', Station_Name,
                      '</br> Income:', Income,
                      '</br> Tract:', tract)) %>%
  layout(title="Median Household Income Along the Red Line [Ashmont/Mattapan]", showlegend=FALSE, xaxis=list(title="", tickangle=-90), yaxis=list(title="", range=c(0, 200000)), margin=list(b=160))
```

####Braintree Red Line
```{r eval = TRUE, echo= FALSE}
mbta = read.csv("mbta_full_red_braintree.csv")
```

```{r eval=TRUE, error=FALSE, message=FALSE, echo=FALSE}
mbta$Station_Name = as.character(mbta$Station_Name)
mbta$Station_Name = factor(mbta$Station_Name, levels=unique(mbta$Station_Name))

plot_ly(mbta, x=~Station_Name, hoverinfo="none") %>%
  add_trace(y=~Income, type="scatter", mode="line", line = list(color="red", width=6)) %>%
  add_trace(y=~Income, type="scatter", mode="markers", marker=list(color="black", size=8), hoverinfo= 'text',
        text = ~paste('</br><b>Station Name:</b>', Station_Name,
                      '</br> Income:', Income,
                      '</br> Tract:', tract)) %>%
  layout(title="Median Household Income Along the Red Line [Braintree]", showlegend=FALSE, xaxis=list(title="", tickangle=-90), yaxis=list(title="", range=c(0, 200000)), margin=list(b=160))
```
*A few notes: *   
*1. Census tracts cover about 1,200 and 8,000 people*  
*2. The census tract for Park St. did not return a median household income, so I used an adjacent census tract instead.*  
*3. The red line splits into 2 lines: one ends in Braintree and the other in Ashmont, hence the 2 graphs.*   
*4. At the Ashmont stop you take the Mattapan Trolley to Mattapan, so I combined these into 1 graph.*  
*5. Sadly, plotly does not support hover on mobile. So, if you want to see more info try hovering over the markers on your laptop.*  
*6. For code details, please see the first [first post](http://translatedmedicine.com/post/income-along-the-boston-t/).*  

**In Part III, we will take a look at the <span style="color:green">The Green Line</span>.**