---
title: The Digital Divide in the U.S.
author: Jorge A Rodriguez MD
date: '2018-01-17'
slug: the-digital-divide-in-the-u-s
---
```{r eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse) 
library(tidycensus) 
library(tilegramsR) 
library(leaflet)
library(leaflet.extras)
library(viridis)


stateNoInt.final = read.csv("stateNoIntFinal.csv")

baseMap = sf_NPR1to1
baseMap.centers = sf_NPR1to1.centers
stateNoInt.final.joined = right_join(baseMap, stateNoInt.final, by="state")
stateNoInt.final.joined.centers = right_join(baseMap.centers, stateNoInt.final, by="state")
popText = paste0(as.character(round(stateNoInt.final.joined$percent, 1)),"%")
getLeafletOptions <- function(minZoom, maxZoom, ...) {
  leafletOptions(
    crs = leafletCRS("L.CRS.Simple"),
    minZoom = minZoom, maxZoom = maxZoom,
    dragging = FALSE, zoomControl = FALSE,
    tap = FALSE,
    attributionControl = FALSE , ...)
}

int.pal = colorBin("plasma", round(stateNoInt.final.joined.centers$percent), pretty=FALSE, 5)

tileGramMap = leaflet(
  stateNoInt.final.joined,
  options= getLeafletOptions(-1.5, -1.5), width = 900) %>%
  addPolygons(
    weight=2,color='#000000', group = 'state',
    fillOpacity = 0.6, opacity = 1, fillColor= ~int.pal(percent),
    highlightOptions = highlightOptions(weight = 4), 
    popup = popText) %>%
  addLabelOnlyMarkers(
    data=stateNoInt.final.joined.centers,
    label = ~as.character(state),
    labelOptions = labelOptions(
      noHide = 'T', textOnly = T,
      offset=c(-4,-10), textsize = '12px')) %>%
  addLegend(pal = int.pal, values = ~percent, opacity = 0.7, position = "bottomright",  
            labFormat = labelFormat(suffix  = "%"), title="Percent of Households Without Internet") %>% 
  setMapWidgetStyle()
```

```{r echo=FALSE}
tileGramMap
```

With the release of the [American Community Survey](https://www.census.gov/programs-surveys/acs/news/data-releases.html), there was an opportunity to take a look at the digital divide in 2016.

I was inspired to make a tilegram map similar to those create by [NPR](http://blog.apps.npr.org/2015/05/11/hex-tile-maps.html). I used the tilegramsR package by [Bhaskar Karambelkar](https://bhaskarvk.github.io/tilegramsR/).

First, here are the packages for the job.
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
#Data clean up
library(tidyverse) 

#Obtaining Census data
library(tidycensus) 

#Needed to create tilegram map
library(tilegramsR) 
library(leaflet)
library(leaflet.extras)
library(viridis)
```

### Census Data
I obtained state level data from the 1 year American Community Survey from 2016:
```{r eval=FALSE, message=FALSE, warning=FALSE}
stateNoInt = get_acs(geography = "state", variables = c("B28002_013E", "B28002_001E"), survey = "acs1")
#B28002_013E --> No Internet Access
#B28002_001E --> Total population
```

I tidy-ed the data so that each row represented one state. I also created the percent column. 
```{r eval=FALSE, message=FALSE, warning=FALSE}
stateNoInt.final= stateNoInt %>% select(-moe) %>%  spread(variable, estimate) %>% 
  mutate(percent = (B28002_013/ B28002_001)*100)
```

The census data uses full state names, while the map data uses state abbreviations. To join the dataframes, I had to match the full state names to the state abbreviations: 
```{r eval=FALSE, message=FALSE, warning=FALSE}
for (i in 1:nrow(stateNoInt.final)) {
  x = stateNoInt.final$NAME[i]
  stateNoInt.final$state[i] = state.abb[match(x,state.name)]
}
```

A few corrections (sorry, Puerto Rico): 
```{r eval=FALSE, message=FALSE, warning=FALSE}
stateNoInt.final$state[9] = "DC"
stateNoInt.final = stateNoInt.final[-52,]
```

```{r eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
stateNoInt.final = read.csv("stateNoIntFinal.csv")
```

### Making the Map
The simple features(sf) maps are provided by the tilegramR package: 
```{r eval=FALSE, message=FALSE, warning=FALSE}
baseMap = sf_NPR1to1
baseMap.centers = sf_NPR1to1.centers
```

Joining the census data with the maps data: 
```{r eval=FALSE, message=FALSE, warning=FALSE}
stateNoInt.final.joined = right_join(baseMap, stateNoInt.final, by="state")
stateNoInt.final.joined.centers = right_join(baseMap.centers, stateNoInt.final, by="state")
```

This part of the code comes from [Rpubs by Bhaskar Karambelkar](http://rpubs.com/bhaskarvk/tilegramsR): 
```{r eval=FALSE, message=FALSE, warning=FALSE}
getLeafletOptions <- function(minZoom, maxZoom, ...) {
  leafletOptions(
    crs = leafletCRS("L.CRS.Simple"),
    minZoom = minZoom, maxZoom = maxZoom,
    dragging = FALSE, zoomControl = FALSE,
    tap = FALSE,
    attributionControl = FALSE , ...)
}
```

I created the bins for the data prior to mapping. The rest of the code is adapted from [Rpubs](http://rpubs.com/bhaskarvk/tilegramsR) and it creates the final map: 
```{r eval=FALSE, message=FALSE, warning=FALSE}
int.pal = colorBin("plasma", round(stateNoInt.final.joined.centers$percent), pretty=FALSE, 5)

tileGramMap = leaflet(
  stateNoInt.final.joined,
  options= getLeafletOptions(-1, -1)) %>%
  addPolygons(
    weight=2,color='#000000', group = 'state',
    fillOpacity = 0.6, opacity = 1, fillColor= ~int.pal(percent),
    highlightOptions = highlightOptions(weight = 4), 
    popup = as.character(round(stateNoInt.final.joined$percent, 1))) %>%
  addLabelOnlyMarkers(
    data=stateNoInt.final.joined.centers,
    label = ~as.character(state),
    labelOptions = labelOptions(
      noHide = 'T', textOnly = T,
      offset=c(-4,-10), textsize = '12px')) %>%
  addLegend(pal = int.pal, values = ~percent, opacity = 0.7, position = "bottomright",  
            labFormat = labelFormat(suffix  = "%"), title="Percent of Households Without Internet") %>% 
  setMapWidgetStyle()
```