---
title: Crimes in Somerville (Time Series)
author: Jorge A Rodriguez MD
date: '2018-01-09'
slug: crimes-in-somerville-i
---

I recently spotted a [time series heat map.](http://aarongonzales.net/posts/nerd/2017/05/18/bpd_cm_advanced/) So I started on a quest to make my own. Luckily, I found some local data to start with: crimes in [Somerville, MA.](https://data.somervillema.gov/Public-Safety/Police-Selected-Criminal-Incidents/4jey-jqxb) This data set includes selected crimes in Somerville from 2005-2017. 


### Handy Packages
```{r message=FALSE}
library(tidyverse)
library(lubridate)
library(ggthemes)
library(plotly)
```


### Getting the Data
```{r}
somerville = read.csv("Police_-_Selected_Criminal_Incidents.csv")
```

### Cleaning the Data
* Remove 2018 data since the year has just started.  
* This part tackles converting the date column into a date data type. Then, I extracted specific parts of the date: the hour, the day of the week, the month, the year, and the day of the month. 
```{r message=FALSE}
somerville$dtreported = mdy_hms(somerville$dtreported)
somerville$wday = wday(somerville$dtreported  , label=TRUE)
somerville$month = month(somerville$dtreported, label=TRUE) #pulls out the month. 
somerville$hour = hour(somerville$dtreported) #pulls out the hour of the day.
somerville$dayomo = mday(somerville$dtreported) #pulls out the day of the month.
somerville$year = year(somerville$dtreported) #pulls out the year. 
somerville = somerville %>%  filter(year !=2018) #removes 2018 data.
```

### Creating the Graph
* I used *dplyr* to to get a count by hour and day of the week.  
* I felt fancy so I used *ggplotly* to make them interactive.
```{r somerville = somerville %>%  filter(year !=2018)}
total = somerville %>% count(hour, wday) %>%  #Getting count by hour and day of the week.
  ggplot(aes(wday, hour, fill=n)) + geom_tile() + #create basic heatmap
  scale_fill_gradient2(guide="legend") + theme_tufte() + #legend structure & tufte theme
  theme(text=element_text(size=14), axis.text = element_text(size=14)) + 
  labs(x="Day of the Week", y="Hour", title="Crimes in Somerville (2005-2017)") + #labels
  guides(fill=guide_legend(title="Crimes")) + 
  scale_y_continuous(breaks=seq(0, 23, 3)) #changing y axis

ggplotly(total)
```

It looks like most crimes happen around 9 in the morning on Tuesdays! 

But what if we look at the breakdown by type of crime. The data set provides an offense variable which has 4 types of crimes: 

* Burglary/breaking and entering 
* Motor vehicle theft 
* Robbery  
* Theft from motor vehicle 

### BURGLARY/BREAKING AND ENTERING
```{r message=FALSE}
burg = somerville %>% filter(offense=="BURGLARY/BREAKING AND ENTERING") %>% 
  count(hour, wday) %>%  #Getting count by hour and day of the week.
  ggplot(aes(wday, hour, fill=n)) + geom_tile() + #create basic heatmap
  scale_fill_gradient2(guide="legend") + theme_tufte() + #legend structure & tufte theme
  theme(text=element_text(size=14), axis.text = element_text(size=14)) + 
  labs(x="Day of the Week", y="Hour", title="Burglary/Breaking and Entering") + #labels
  guides(fill=guide_legend(title="Crimes")) + 
  scale_y_continuous(breaks=seq(0, 23, 3)) #changing y axis

ggplotly(burg)
```

It seems that most burglaries happen on Wednesday evenings.

### MOTOR VEHICLE THEFT 
```{r message=FALSE}
carTheft = somerville %>% filter(offense=="MOTOR VEHICLE THEFT") %>% 
  count(hour, wday) %>%  #Getting count by hour and day of the week.
  ggplot(aes(wday, hour, fill=n)) + geom_tile() + #create basic heatmap
  scale_fill_gradient2(guide="legend") + theme_tufte() + #legend structure & tufte theme
  theme(text=element_text(size=14), axis.text = element_text(size=14)) + 
  labs(x="Day of the Week", y="Hour", title="Motor Vehicle Theft") + #labels
  guides(fill=guide_legend(title="Crimes")) + 
  scale_y_continuous(breaks=seq(0, 23, 3)) #changing y axis

ggplotly(carTheft)
```

Car thefts seem to happen on Monday afternoons and around noon time on Thursdays.

### ROBBERY
```{r message=FALSE}
rob = somerville %>% filter(offense=="ROBBERY") %>% 
  count(hour, wday) %>%  #Getting count by hour and day of the week.
  ggplot(aes(wday, hour, fill=n)) + geom_tile() + #create basic heatmap
  scale_fill_gradient2(guide="legend") + theme_tufte() + #legend structure & tufte theme
  theme(text=element_text(size=14), axis.text = element_text(size=14)) + 
  labs(x="Day of the Week", y="Hour", title="Robbery") + #labels
  guides(fill=guide_legend(title="Crimes")) + 
  scale_y_continuous(breaks=seq(0, 23, 3)) #changing y axis

ggplotly(rob)
```

This makes sense! Most robberies happen Saturday late at night/early in the morning.

### THEFT FROM MOTOR VEHICLE
```{r message=FALSE}
theft = somerville %>% filter(offense=="THEFT FROM MOTOR VEHICLE") %>% 
  count(hour, wday) %>%  #Getting count by hour and day of the week.
  ggplot(aes(wday, hour, fill=n)) + geom_tile() + #create basic heatmap
  scale_fill_gradient2(guide="legend") + theme_tufte() + #legend structure & tufte theme
  theme(text=element_text(size=14), axis.text = element_text(size=14)) + 
  labs(x="Day of the Week", y="Hour", title="Theft from Motor Vehicle") + #labels
  guides(fill=guide_legend(title="Crimes")) + 
  scale_y_continuous(breaks=seq(0, 23, 3)) #changing y axis

#Making it interactive
ggplotly(theft)
```
It looks like most thefts from motor vehicles happen in the morning on weekdays.

There you have it a quick time series visualiation of crime in Somerville.
