---
title: Income Along the Boston T [Blue Line]
author: Jorge A Rodriguez MD
date: '2018-01-04'
slug: income-along-the-boston-t-iv
---

In the first three parts of this series, we looked at the changes in income along the [<span style="color:orange">The Orange Line</span>](http://translatedmedicine.com/post/income-along-the-boston-t/), [<span style="color:red">The Red Line</span>](http://translatedmedicine.com/post/income-along-the-boston-t-ii/), and [<span style="color:green">The Green Line</span>](http://translatedmedicine.com/post/income-along-the-boston-t-iii/). I went through [my approach](http://translatedmedicine.com/post/income-along-the-boston-t/) for creating a [New Yorker](https://projects.newyorker.com/story/subway/) inspired visualization of income along the Boston T. 

For the **final post**, we take a look at the <span style="color:blue">The Blue Line</span>. Below you will see median household income along each of the four lines of the Boston Blue Line.

####Blue Line
```{r eval = TRUE, echo= FALSE, message=FALSE}
library(plotly)
mbta = read.csv("mbta_full_blue.csv")
```

```{r eval=TRUE, error=FALSE, message=FALSE, echo=FALSE}
mbta$Station_Name = as.character(mbta$Station_Name)
mbta$Station_Name = factor(mbta$Station_Name, levels=unique(mbta$Station_Name))

plot_ly(mbta, x=~Station_Name, hoverinfo="none") %>%
  add_trace(y=~Income, type="scatter", mode="line", line = list(color="blue", width=6)) %>%
  add_trace(y=~Income, type="scatter", mode="markers", marker=list(color="black", size=8), hoverinfo= 'text',
        text = ~paste('</br><b>Station Name:</b>', Station_Name,
                      '</br> Income:', Income,
                      '</br> Tract:', tract)) %>%
  layout(title="Median Household Income Along the Blue Line", showlegend=FALSE, xaxis=list(title="", tickangle=-90), yaxis=list(title="", range=c(0, 200000)), margin=list(b=160))
```




*A few notes: *   
*1. Census tracts cover about 1,200 and 8,000 people.*  
*2. Sadly, plotly does not support hover on mobile. So, if you want to see more info try hovering over the markers on your laptop.*  
*3. For code details, please see the first [first post](http://translatedmedicine.com/post/income-along-the-boston-t/).*  


