---
title: "Where are the students who are English Learners in MA?"
author: "Jorge A Rodriguez MD"
date: '2018-04-29'
slug: where-are-the-students-who-are-english-learners-in-ma
---
Inspired by my wife's work in educating students who are English Learners, this post visualizes the English Learners across Massachusetts.

The Massachusetts Department of Education publishes a large amount of education data from across the state. It provides yearly percentages of enrolled student and their characteristics, including English Learners. English Learners are defined as students age 3-21 who were not born in the US or whose native language is not English. They suffer educational disparities and have higher dropout rates than their English speaking counterparts.

With that introduction in mind, lets get to it.

# The packages: 
```{r warning=FALSE, eval=TRUE, error= FALSE, message=FALSE}
library(tidyverse) # a classic
library(viridis) # for color palettes 
library(sf) #to read a display shapefiles that create the map
library(leaflet) # to make interactive map
```

# The data: 
## Source
The map file comes from [Mass GIS](https://docs.digital.mass.gov/dataset/massgis-data-public-school-districts).
The education data from the [Massachusetts Department of Education](http://profiles.doe.mass.edu/state_report/selectedpopulations.aspx?mode=district&year=2018&orderBy=LEPPCT%20DESC) 
The **district_sf** variable contains the shapefile for the map. The **elma** variable contains the data on English Learners in each district. 
```{r warning=FALSE, eval=TRUE, error= FALSE, message=FALSE}
district_sf = st_read("schooldistricts/SCHOOLDISTRICTS_POLY.shp")
elma = read.csv("selectedpopulations.xls.csv", stringsAsFactors = F, header = T)
elma = elma[-1,] #removes an empty row
```

# The wrangling: 
## Step 1
There were a few variable types that needed to be changed. The **paste0** creates a new variable with a leading zero for district code to help in merging data sets later on.
```{r warning=FALSE, eval=TRUE, error= FALSE, message=FALSE}
district_sf$ORG_CODE = as.character(district_sf$ORG_CODE) 
elma$ORGCODE = as.character(elma$ORGCODE)

district_sf$DISTRICT = as.character(district_sf$DISTRICT)
elma$ï..DISTRICT = as.character(elma$ï..DISTRICT)

elma$ORGCODE_E = paste0("0", elma$ORGCODE)
```

## Step 2
This was by far the trickiest step! I use a series of left joins to join the data from the MA DOE to the district level shapefile. The challenge is was that after each left join I had a column with some rows with NAs. In order to combine the columns and fill in the NAs, I used the  **mutate** step used the **coaslece** function. My understanding is that the **coaslece** function combines columns and replaces rows with NA values with from other columns that have a value listed for that row.
```{r warning=FALSE, eval=TRUE, error= FALSE, message=FALSE}
district_sf = district_sf %>% left_join(elma, by=c("ORG_CODE" = "ORGCODE")) %>%
  left_join(elma, by=c("ORG_CODE" = "ORGCODE_E")) %>%
  left_join(elma, by = c("DISTRICT"="ï..DISTRICT")) %>%
  mutate(XEL1 = coalesce(X.1.x, X.1.y, X.1))

district_sf$XEL1 = as.numeric(district_sf$XEL1)
```

##Step 3
I removed the charter districts since they covered large areas and were different than the rest of the districts listed.
```{r warning=FALSE, eval=TRUE, error= FALSE, message=FALSE}
district_sf = district_sf %>% filter(MADISTTYPE != "Charter - Commonwealth" & MADISTTYPE != "Charter - Horace Mann")
```

#The map: 
The map represents the percentage of EL students in each district presented as percentiles.
```{r warning=FALSE, eval=TRUE, error= FALSE, message=FALSE}
pal = colorQuantile(palette = "viridis", domain = as.numeric(district_sf$XEL1), n = 10)

mapel = district_sf %>%
  st_transform(crs = '+proj=longlat +datum=WGS84') %>% 
  leaflet(width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(popup = ~ paste0(DISTRICT, ":", XEL1, "%"),
              stroke = FALSE,
              smoothFactor = 0,
              weight = 2,
              fillOpacity = 0.7, 
              color = ~ pal(XEL1)) %>% 
  addLegend("bottomright", 
            pal = pal, 
            values = ~ XEL1,
            title = "% English Learners in MA (Percentiles)",
            opacity = 1)

mapel
```

That's it! The MA Department of Education has data dating back to the 1990s, so my next step is to tackle these years with the hopes of creating a Shiny app to explore the data over time.